File-based localization of numerical perturbations in data analysis pipelines

### 2020-03-27 paper.tex
reply comments in results

The HCP developed a set of pre-processing pipelines to process structural,
functional, and diffusion MRI data acquired in the project. We focus on HCP
pre-processing pipelines for structural data, and particularly
PreFreeSurfer and FreeSurfer. 
A detailed description of the analyses done in these
pipelines is available in~\cite{glasser2013}. 
In summary, PreFreeSurfer consists of the following steps: 
\begin{itemize}
\item Distortion Correction (DC), 
\item Anatomical Average (AAve), 
\item ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A), 
\item Brain Extraction (BExt), 
\item Bias Field Correction (BFC), 
\item Atlas-Registration (AR).
\end{itemize}
And FreeSurfer consists of the following ones:
\begin{itemize}
\item Downsampling images, 
\item T1w image registration, 
\item Placement of surface images, 
\item Folding-based surface registration.
\end{itemize}
The average processing time per subject is approximately 2 hours for
PreFreeSurfer and 8 hours for FreeSurfer. The average output file size is
approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
\tristan{Ali, are these numbers per subject or for 20 subjects?}

We randomly selected 20 unprocessed subjects 
from the HCP data release S500 
available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
\tristan{Can you let me know where the data is located, so that I could describe it here?}
See~\cite{van2013wu} for a discussion of these various modalities.

@@

\tristan{See what to do with this paragraph}
Within-OS variability may also be coming from other factors than
pseudo-randomness. For instance, the same files may have different metadata
such as timestamps, file paths, and other execution-specific information.
To ensure that the differences are not related to the metadata, we used
specific functions from FreeSurfer tool including \emph{mri\_diff},
\emph{mris\_diff}, and \emph{lta\_diff}. These functions compare files
based on the image dimensions, voxel intensities, the number of vertices,
and geometry information (see~\cite{fischl2012freesurfer} for more
details) \tristan{Ali, are these functions used by \toolname to compare results?}.

@@

% make figure in two columns use {figure*}
\begin{figure*}
\centering
  \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
  \caption{Frequency heatmap of non-reproducible PreFreeSurfer processes (N=20). 
           Rows refer to PreFreeSurfer pipeline steps.
           Cell labels describe the frequency distribution of the instances as ``N [M] x Y" 
           in which N is the proportion of the process 
           that creates differences out of all M times that is occurred. 
           Also, times Y shows the number of processes with the same distribution.
           For example, the cell of FLIRT tool in the sub-pipeline AAve shows 10 instances of the process 
           with the frequency of 11 that introduce no differences.
           The range between red and green colors shows the highest and lowest 
           proportion of occurrences of the processes with differences.
           Also, grey cells indicate that the tool is not executed in the sub-pipeline.}
  \label{fig:pfs_freq}
\end{figure*}

To give an idea of the observed differences, Figure~\ref{fig:fnirt_result} 
compares results of ACPC-Alignment in PreFreeSurfer. \tristan{Ali, there is no
FNIRT process in ACPC-Alignment, please check.}

@@

The only non-reproducible process identified in FreeSurfer was
\texttt{mris\_make\_surfaces}, which was the case for 10 out of 20
subjects. \tristan{do you know why this was the case: is this CLI
dynamically linked?} This process generates the cortical and white matter
surfaces. However, this does not mean that FreeSurfer results were
identical for the remaining 10 subjects and for the other processes, due to the
propagation of PreFreeSurfer differences.

>>>>

The HCP developed a set of pre-processing pipelines to process structural,
functional, and diffusion MRI data acquired in the project. We focus on HCP
pre-processing pipelines for structural data, and particularly
PreFreeSurfer and FreeSurfer. 
A detailed description of the analyses done in these
pipelines is available in~\cite{glasser2013}. 
In summary, PreFreeSurfer consists of the following steps: 
\begin{itemize}
\item Distortion Correction (DC), 
\item Anatomical Average (AAve), 
\item ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A), 
\item Brain Extraction (BExt), 
\item Bias Field Correction (BFC), 
\item Atlas-Registration (AR).
\end{itemize}
And FreeSurfer consists of the following ones:
\begin{itemize}
\item Downsampling images, 
\item T1w image registration, 
\item Placement of surface images, 
\item Folding-based surface registration.
\end{itemize}
The average processing time per subject is approximately 2 hours for
PreFreeSurfer and 8 hours for FreeSurfer. The average output file size is
approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}

We randomly selected 20 unprocessed subjects 
from the HCP data release S500 
available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
\tristan{Can you let me know where the data is located, so that I could describe it here?
Subjects are located in consider:/data/asalari/ali-tests/PFS\_Centos6\_Traced
and consider:/data/asalari/ali-tests/PFS\_Centos7\_Traced 
and outputs are in consider:/data/asalari/ali-tests/nurm-out}
See~\cite{van2013wu} for a discussion of these various modalities.

@@

\tristan{See what to do with this paragraph}
Within-OS variability may also be coming from other factors than
pseudo-randomness. For instance, the same files may have different metadata
such as timestamps, file paths, and other execution-specific information.
To ensure that the differences are not related to the metadata, we used
specific functions from FreeSurfer tool including \emph{mri\_diff},
\emph{mris\_diff}, and \emph{lta\_diff}. These functions compare files
based on the image dimensions, voxel intensities, the number of vertices,
and geometry information (see~\cite{fischl2012freesurfer} for more
details) \tristan{Ali, are these functions used by \toolname to compare results?
Yes, they have been used based on the file format}.

@@

% make figure in two columns use {figure*}
\begin{figure*}
\centering
  \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
  \caption{Frequency heatmap of non-reproducible PreFreeSurfer processes (N=20). 
           Rows refer to PreFreeSurfer pipeline steps.
           Cell labels describe the frequency distribution of the instances as ``N [M] x Y" 
           in which N is the proportion of the process 
           that creates differences out of all M times that is occurred. 
           Also, times Y shows the number of processes with the same distribution.
           For example, the FNIRT process occurred at most 6 times in sub-pipeline DC in each subject. 
           Also, this shows 3 different distributions among the subjects.
           The left cell shows 2 instances of the FNIRT that introduce differences in 19 out of 20 subjects.
           The middle cell indicates three instances of the FNIRT that create no differences for all 20 subjects.
           The right cell shows one instance of the FNIRT that introduces differences in 17 out of 20 subjects.
           The range between red and green colors shows the highest and lowest 
           proportion of occurrences of the processes with differences.
           Also, grey cells indicate that the tool is not executed in the sub-pipeline.}
  \label{fig:pfs_freq}
\end{figure*}

To give an idea of the observed differences, Figure~\ref{fig:fnirt_result} 
compares results of Brain Extraction in PreFreeSurfer. \tristan{Ali, there is no
FNIRT process in ACPC-Alignment, please check. Ali: yes, it was brain extraction.}

@@

The only non-reproducible process identified in FreeSurfer was
\texttt{mris\_make\_surfaces}, which was the case for 10 out of 20
subjects. \tristan{do you know why this was the case: is this CLI
dynamically linked? Ali: It includes dynamic linking and I guess it may be affected by this.} 
This process generates the cortical and white matter
surfaces. However, this does not mean that FreeSurfer results were
identical for the remaining 10 subjects and for the other processes, due to the
propagation of PreFreeSurfer differences.
